Type Refinements for the Working Class

### 2015-11-13 paper.tex
Formatting
Use tikz for inline arrows

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{leftidx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{catchfilebetweentags}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{ifdraft}
\usepackage{ifthen}
\usepackage{mathtools}
\usepackage{proof}
\usepackage{scalerel}
\usepackage{setspace}
\usepackage{stmaryrd}
\usepackage{supertabular}
\usepackage{tikz-cd}
\usepackage{url}
\usepackage{todonotes}

@@

\newcommand\Cats{\mathbf{Cat}}
\newcommand\Discrete[1]{#1_{\equiv}}
\newcommand\Arr[1]{\mathbf{{#1}^{\rightarrow}}}
\newcommand\Sets{\mathbf{Set}}
\newcommand\SCtx{\mathbf{SCtx}}
\newcommand\Ctx{\mathbf{Ctx}}
\newcommand\Seq{\mathbf{Seq}}
\newcommand\HypCat{\mathbf{H}}

@@

For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
a generic logical framework. Recall from~\cite{sterling-morrison:2015} that
$\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
renamings. Then, fix a small category $\mathcal{W}$ of worlds along with a map
$\Of{\pi}{\mathcal{W}\to\SCtx}$ such that $\pi$ is a Grothendieck
fibration~\cite{jacobs:2001, taylor:1999}:
\[
  \begin{tikzcd}[cramped, ampersand replacement = \&, row sep = large]
    {\IMode{\mathcal{W}}}
      \arrow[dash, d]
\&
    {\IMode{{w}_{i}}}
      \arrow[drrr]
      \arrow[dotted, dr]
      \arrow[mapsto, dd]
\&
    \null
\&
    \null
\&
    \null
\\
    {\IMode{{\pi}}}
      \arrow[d]
\&
    \null
\&
    {\OMode{{w}_{j}}}
      \arrow[rr, swap, "\OMode{{\iota}_{\varrho}}"]
      \arrow[mapsto, dd, bend right = 50]
      \arrow[mapsto, d]
\&
    \null
\&
    |[alias = Wk]|{\IMode{{w}_{k}}}
      \arrow[mapsto, d]
\\
    {\IMode{\SCtx}}
\&
    |[alias = Ui]|{\IMode{{\Upsilon}_{i}}}
      \arrow[dr]
\&
    {\OMode{\scriptstyle{{\pi}({w}_{j})}}}
      \arrow[rr, "\OMode{{\pi}({\iota}_{\varrho})}", ""{name = Msrc, swap}]
      \arrow[equals, d]
\&
    \null
\&
    {\IMode{\scriptstyle{{\pi}({w}_{k})}}}
      \arrow[equals, d]
\\
    \null
\&
    \null
\&
    {\IMode{{\Upsilon}_{j}}}
      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]\&
    \null
\&
    |[alias = Uk]|{\IMode{{\Upsilon}_{k}}}
      \arrow[mapsto, from = Wk, to = Uk, bend right = 50, crossing over]
      \arrow[equals, from = Msrc, to = Mtgt]
      \arrow[from = Ui, to = Uk, crossing over]
  \end{tikzcd}
\]
Then, a \emph{judgment} shall be interpreted as an object in the presheaf
category $\Define{\mathbb{J}}{\Sets^{\OpCat{\mathcal{W}}}}$; in other words, a
judgment is identified with the intensional set of its renaming-sensitive
derivations.

@@

  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Omega}}$, we can define
  the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows, where
  $\Of{\Yoneda{-}}{\mathcal{W}\to\mathbb{J}}$ is the Yoneda Embedding:
  \begin{align*}
    \ADefine{
      \left(\GenJ{\Omega}{\mathcal{J}}\right)(w)
    }{
      \Hom{\mathbb{J}}{
        \Yoneda{w}
      }{
        Ran_{\mathbbold{1}}{\mathcal{J}}
      }
    }\\
    &\cong
    \OMode{
      \int_{(w',\,\vec{E})}{{w'}\succeq{w}\Rightarrow{\mathcal{J}(\vec{E})(w')}}
    }
  \end{align*}

@@

The second judgment is defined coinductively via the following meaning
explanation:
\begin{definition}[Parametric Refinement]
  To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
  $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
  $\Of{\varrho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, and for any $M, N$,
  what it means for $M$ and $N$ to be identified under $\phi$, supposing
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
  $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
  $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
  $\IMode{M}\sim^\rho_\phi\IMode{O}$.
\end{definition}

@@

\todo[inline]{%
  define maps of refined contexts, then change weakening lemma to
  monotonicity, which is stronger and more useful.
}
\begin{thm}[Weakening]
  If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
  $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
\end{thm}
\begin{proof}
  This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
  $\Upsilon\hookrightarrow\Upsilon,u:\sigma$.
\end{proof}

@@

Two refinements are equal when they denote the same PER.\@That is, will define
the refinement equality judgment $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ as
follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$:
\begin{definition}
  To know $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
  $\Of{\rho}{\Upsilon\hookrightarrow\OMode{\Upsilon'}}$, the following judgments:
  \begin{gather*}
    \GenJ{
      \MV{m},\MV{n}:\tau
    }{
      \HypJ{
        \IMode{M}\sim^\rho_\phi\IMode{N}
      }{
        \IMode{M}\sim^\rho_\psi\IMode{N}
      }
    }\\
    \GenJ{
      \MV{m},\MV{n}:\tau
    }{
      \HypJ{
        \IMode{M}\sim^\rho_\psi\IMode{N}
      }{
        \IMode{M}\sim^\rho_\phi\IMode{N}
      }
    }
  \end{gather*}

@@

The primary judgment concerning refinements is the parametric equality,
$\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
$\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
$\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
$M$ and $N$ are identified by $\phi$ at the identity renaming
$\Of{1}{\IMode{\Upsilon}\hookrightarrow\OMode{\Upsilon}}$:
\[
  \infer{
    \RMemEq{\Phi}{M}{N}{\phi}
  }{
    \IMode{M}\sim^1_\phi\IMode{N}
  }
\]

>>>>

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{leftidx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{catchfilebetweentags}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{ifdraft}
\usepackage{ifthen}
\usepackage{mathtools}
\usepackage{proof}
\usepackage{scalerel}
\usepackage{setspace}
\usepackage{stmaryrd}
\usepackage{supertabular}
\usepackage{tikz-cd}
\usepackage{tikzpfeile}
\usepackage{url}
\usepackage{todonotes}

@@

\newcommand\Cats{\mathbf{Cat}}
\newcommand\Discrete[1]{#1_{\equiv}}
\newcommand\Arr[1]{\mathbf{{#1}^{\ra}}}
\newcommand\Sets{\mathbf{Set}}
\newcommand\SCtx{\mathbf{SCtx}}
\newcommand\Ctx{\mathbf{Ctx}}
\newcommand\Seq{\mathbf{Seq}}
\newcommand\HypCat{\mathbf{H}}

@@

For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
a generic logical framework. Recall from~\cite{sterling-morrison:2015} that
$\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
renamings. Then, fix a small category $\mathcal{W}$ of worlds along with a map
$\Of{\pi}{\mathcal{W}\ra\SCtx}$ such that $\pi$ is a Grothendieck
fibration~\cite{jacobs:2001, taylor:1999}:
\[
  \begin{tikzcd}[cramped, ampersand replacement = \&, row sep = large]
    {\IMode{\mathcal{W}}}
      \arrow[dash, d]
\&
    {\IMode{{w}_{i}}}
      \arrow[drrr]
      \arrow[dotted, dr]
      \arrow[mapsto, dd]
\&
    \null
\&
    \null
\&
    \null
\\
    {\IMode{{\pi}}}
      \arrow[d]
\&
    \null
\&
    {\OMode{{w}_{j}}}
      \arrow[rr, swap, "\OMode{{\iota}_{\varrho}}"]
      \arrow[mapsto, dd, bend right = 50]
      \arrow[mapsto, d]
\&
    \null
\&
    |[alias = Wk]|{\IMode{{w}_{k}}}
      \arrow[mapsto, d]
\\
    {\IMode{\SCtx}}
\&
    |[alias = Ui]|{\IMode{{\Upsilon}_{i}}}
      \arrow[dr]
\&
    {\OMode{\scriptstyle{{\pi}({w}_{j})}}}
      \arrow[rr, "\OMode{{\pi}({\iota}_{\varrho})}", ""{name = Msrc, swap}]
      \arrow[equals, d]
\&
    \null
\&
    {\IMode{\scriptstyle{{\pi}({w}_{k})}}}
      \arrow[equals, d]
\\
    \null
\&
    \null
\&
    {\IMode{{\Upsilon}_{j}}}
      \arrow[rr, swap, "{\IMode{\varrho}}", ""{name = Mtgt, swap}]
\&
    \null
\&
    |[alias = Uk]|{\IMode{{\Upsilon}_{k}}}
      \arrow[mapsto, from = Wk, to = Uk, bend right = 50, crossing over]
      \arrow[equals, from = Msrc, to = Mtgt]
      \arrow[from = Ui, to = Uk, crossing over]
  \end{tikzcd}
\]
Then, a \emph{judgment} shall be interpreted as an object in the presheaf
category $\Define{\mathbb{J}}{\Sets^{\OpCat{\mathcal{W}}}}$; in other words, a
judgment is identified with the intensional set of its renaming-sensitive
derivations.

@@

  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Omega}}$, we can define
  the general judgment $\GenJ{\Omega}{\mathcal{J}}$ as follows, where
  $\Of{\Yoneda{-}}{\mathcal{W}\ra\mathbb{J}}$ is the Yoneda Embedding:
  \begin{align*}
    \ADefine{
      \left(\GenJ{\Omega}{\mathcal{J}}\right)(w)
    }{
      \Hom{\mathbb{J}}{
        \Yoneda{w}
      }{
        Ran_{\mathbbold{1}}{\mathcal{J}}
      }
    }\\
    &\cong
    \OMode{
      \int_{(w',\,\vec{E})}{{w'}\succeq{w}\Rightarrow{\mathcal{J}(\vec{E})(w')}}
    }
  \end{align*}

@@

The second judgment is defined coinductively via the following meaning
explanation:
\begin{definition}[Parametric Refinement]
  To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
  $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
  $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
  what it means for $M$ and $N$ to be identified under $\phi$, supposing
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
  $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
  $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
  $\IMode{M}\sim^\rho_\phi\IMode{O}$.
\end{definition}

@@

\todo[inline]{%
  define maps of refined contexts, then change weakening lemma to
  monotonicity, which is stronger and more useful.
}
\begin{thm}[Weakening]
  If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
  $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
\end{thm}
\begin{proof}
  This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
  $\Upsilon\inj\Upsilon,u:\sigma$.
\end{proof}

@@

Two refinements are equal when they denote the same PER.\@That is, will define
the refinement equality judgment $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ as
follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$:
\begin{definition}
  To know $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
  $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
  \begin{gather*}
    \GenJ{
      \MV{m},\MV{n}:\tau
    }{
      \HypJ{
        \IMode{M}\sim^\rho_\phi\IMode{N}
      }{
        \IMode{M}\sim^\rho_\psi\IMode{N}
      }
    }\\
    \GenJ{
      \MV{m},\MV{n}:\tau
    }{
      \HypJ{
        \IMode{M}\sim^\rho_\psi\IMode{N}
      }{
        \IMode{M}\sim^\rho_\phi\IMode{N}
      }
    }
  \end{gather*}

@@

The primary judgment concerning refinements is the parametric equality,
$\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
$\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
$\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
$M$ and $N$ are identified by $\phi$ at the identity renaming
$\Of{1}{\IMode{\Upsilon}\inj\OMode{\Upsilon}}$:
\[
  \infer{
    \RMemEq{\Phi}{M}{N}{\phi}
  }{
    \IMode{M}\sim^1_\phi\IMode{N}
  }
\]
